---
title: "lab_02"
author: "derek willis"
date: "8/09/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings
```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
library(lubridate)
```

## Load Data

Let's keep working with the slice of WinRed contribution data from West Virginia.  

```{r}
wv_winred_contribs <-read_rds("data/lab_02.rds")
```

## Examine Data

## Answer questions

**Q1** Create a new column that divides the contribution amounts into the following options - less than $10, $10-$25, $26-$50, $51_$100, $101-$500, $501 and up - and calculates how many contributions are in each group. How would you describe the distribution of contributions in a sentence?
A1. 
 The first section, there's a $15 difference. The second section there's a $25 difference, then a $50 difference, and the last one is a $400 difference.
```{r}
  contrib_amounts <- wv_winred_contribs %>%
  mutate(
    amount_category = case_when(
        amount < 10 ~ "Less than $10",
        amount >= 10 & amount <= 25 ~ "Between $10 and $25",
        amount > 26 & amount <= 50 ~ "Between $26 and $50",
        amount > 51 & amount <= 100 ~ "Between $51 and $100",
        amount > 101 & amount <= 500 ~ "Between $101 and $500",
        amount > 501 ~ "More than $501"
    )
)

```

Q2. Which West Virginia donors made the most contributions of less than $10? You'll need to use a combination of columns to get unique donors.  
A2. 

```{r}
contrib_amounts %>% 
  group_by(amount_category) %>% 
  summarize(total_amount = sum(amount)) %>% 
  arrange(desc(total_amount))
``` 


```{r}
 less_than_10 <- wv_winred_contribs %>% 
  filter (amount<10) %>% 
  group_by (first_name,last_name) %>%  
  summarise (number_of_contributions=n()) %>%  
  arrange (desc(number_of_contributions))
less_than_10
```



Q3. What questions do the results of Q2 raise in your mind? How could you use the data to explore them? No code here, just your questions.

A3. I could use why specific donors are related to organization they chose to donate to. 

Q4. Which month saw the most contributions?
A4. 

```{r}
wv_winred_contribs %>% 
  mutate (month= floor_date(date,"month")) %>% 
group_by(month) %>%  
    summarise(total=n())) %>%  
    arrange(desc(total))


```

Q5. How many contributions of $2,900 - the individual maximum donation - were there, and who did they go to? How many of the recipients are from West Virginia? Do some googling to find out.

A5.I don't know.  
I'm really frustrated because now its saying it's not reading the "read.rsv.." So maybe nothing is coming out right because of the very first step but It was fine when I was in class and saved it. Even Thursday, I was working on it and the "rsv" wasn't  a problem until now.  I know I can get it. This time, I just have to submit what I have,and I think it's all I can think of 
```{r}
wv_winred_contribs %>% 
  filter(amount=2900) %>%  
  arrange (date)
```
