---
title: "lab_01"
author: "Derek Willis"
date: "7/21/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Turn off scientific notation
options(scipen=999)
```

## About this lab

To complete this lab, you need to:
* run existing code as directed (look for **Task**).
* modify existing code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).
When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

Remember to add any new verbs or concepts to your reference notebook!

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
install.packages('tidyverse')
library(tidyverse)

```

## Load Data

We're going to load the slice of WinRed contribution data from West Virginia.

**Task** Run the codeblock below to load the data, which is contained in an RDS file.

```{r}
wv_winred_contribs <- read_rds("lab_01.rds")
```

## Examine Data

**Task** Run the codeblock below to use glimpse to get a sense of the column names, types and sample data.
**Q1** How many rows and columns does this dataframe have?
**A1** 
This dataframe has 13,382 rows and 24 columns. 
```{r}
glimpse(wv_winred_contribs)
```

## Answer questions

**Q2** Which West Virginia city provided the largest number of contributions? How many?
**A2** 
Martinsburg had the largest number of contributions.I panicked and thought counting the contributions was under the variable "amount" but it's the same variable name from the last assignment. Yay for problem-solving skills that should've came sooner. A lot of the code is me going back from the beginning to figure it out. 

```{r} 
wv_winred_contribs %>% 
  group_by(city) %>% 
  summarise(
count_contribs= n()
) %>% 
arrange(desc(count_contribs))
```
**Q3** Which West Virginia city accounted for the fewest contributions? How many?
**A3** So I tried to run the fewest contributions in the codeblock before. For some reason, everytime I tried to do it in this codeblock, the error would say something about zero-length variable. I can only see the fewest contributions if I do it in the codeblock above. The fewest contributions are from the following cities: Auburn Belle, Belmont, Benwood, Camden on Gauley, Cedarville, Ceredo, Coldwater Creek, Colliers, Crawley,Cyclone, Dryfork, Dunmore, Durbin, Ellenboro, Evans, Exchange, Foster, Gordon, Huttonsville,Independence, Jeffery, Jolo, Mason, Matheny, Midway, Mill Creek, Mount Storm, New Milton, Newburg, Newell, Ranger, Saint ALBANS, Shoals, Smoot, Sophia, Springfield, Surveyor, Troy, and Unger. All of their contributions were one.   
 
```{r}
wv_winred_contribs %>% 
  group_by(city) %>% 
  summarise(
count_contribs= n()
) %>% 
arrange((count_contribs))
``` 
**Q4** What's your best educated guess to explain why the city identified in question 2 had more contributions compared to the cities identified in question 3? Hint: you won't find the answer in the data, you'll need to do some quick web research. You do not need to write any code for this question, so no codeblock is provided.  

**A4** I believe it's because the cities are significantly smaller or have low population for the cities that have the lowest contribution. 

**Q5** What recipient got the largest total amount of money? How many contributions did they receive and what was the average donation?

**A5** 
The recipient that had the largest total amount of money was Save America Joint Fundraising Committee. Their total amount of money was 66972.89. The average contribution was 27.35820670, rounded to 27.4 or 27 contributions 
```{r}
wv_winred_contribs %>%
  group_by(memo_text) %>%
  summarise(
    count_contribs = n(),
    total_amount = sum(amount),
    mean_amount = mean(amount),
  ) %>%
  arrange(desc(count_contribs))
```

**Q6** What is that top recipient and what explains its popularity? Hint: this will require web research, you won't be able to answer this purely from the data set.  You do not need to write any code for this question, so no codeblock is provided.  
**A6** The top recipient is Save America Joint Fundraising Committee. I see now, it's popularity is because it's Donald Trump's Official Leadership PAC. His name is attached to the Fundraising Committee where the NRSC and the Republican National Committee is very vague and people gravitate if it's one specifc person they recognize.  

**Q7**. Produce a summary table that has the following information for West Virginia as a whole (the whole state, not by city):
* count of contributions
* total amount of all contributions
* the average contribution amount
* the median contribution amount
* the smallest contribution amount
* the largest contribution amount

Store it as a variable object called `wv_summary` so that you can examine it with the data viewer. What newsworthy nuggets jump out at you that you could imagine dropping into a news story? Write a few summary sentences with overall findings.

**A7** Some newsworthy points would be how the size of a city impacts the amount of contributions made. I defintely could mention how most of the recipients that did really well used their name in the committee and I believed it helped people to donate more money with the exception of the Republican National Committee 

```{r}
wv_winred_contribs %>%
  group_by(city) %>%
  summarise(
    count_contribs = n(),
    total_amount = sum(amount),
    mean_amount = mean(amount),
    median_amount = median(amount),
    min_amount = min(amount),
    max_amount = max(amount)
  ) %>%
  arrange(desc(total_amount))
wv_winred_contribs -> wv_summary
```
