---
title: "lab_07"
author: "derek willis"
date: "10/21/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

* A Census API key

## Load libraries and establish settings

You'll need to load two packages for this: the tidyverse and tidycensus.

**Task** load these two packages

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the libraries you'll need
library(tidyverse)
library(tidycensus)
```

## Setup Census API

You'll need your Census API Key:

**Task** set your API Key if you haven't already

```{r echo=FALSE}
census_api_key("cb528bbc0c6c5be1ab0e86b0029baed11efff462", overwrite=TRUE)
```

## Load and view ACS variables

You'll also need to load the list of American Community Survey variables from the 2020 5-year estimates so you can identify the codes you need:

**Task** load the variables from the 2020 5-year ACS and save them to a dataframe

```{r}
acs5 <- load_variables(2020, "acs5") 

median20<- get_acs(geography = "county", variables= "B01002_001", year= 2020, state= 'MD')
```

## Answer questions

**Q1** What is the Maryland jurisdiction with the lowest median age, according to the ACS 2020 5-year estimates? You will need to find the variable for median age first.
**A1**
Baltimore City, Maryland has the lowest median age as 35.5 years old. 
```{r}
median20<- get_acs(geography = "county", variables= "B01002_001", year= 2020, state= 'MD') %>% 

arrange(estimate)

```

**Q2** Which Maryland counties have a median age of at least 50? You can use the dataframe you just made to answer this, but you must write code to display only those counties with a median age of at least 50.
**A2** 
Worcester County and Talbot County
```{r}
median20 %>%  
   mutate(
   med50 = case_when(
        estimate >=50 ~ "at least age of 50" )
  ) %>% 
  
  arrange(desc(med50))


```

**Q3** We're interested in knowing which Maryland counties have the highest percentage of 5-17 year-olds who speak Spanish at home and *also* speak English fluently. Start with the code you wrote for Q1 and alter it to show the total population of 5-17 year-olds by county. Then create a second dataframe to show the number of 5-17 year-olds by county living in homes where Spanish is spoken but who speak English "very well" according to the ACS. To do this, you will need to find the ACS variables that represent those subjects. HINT: the phrase "very well" shows up in multiple variables - you want the one that doesn't include other factors like citizenship. 

For each dataframe, rename the `estimate` column to a meaningful name you've given the ACS variable, remove the `variable` columns in each using select and then join them on GEOID and NAME. Then you can add a column to calculate the percentage of the total 5-17 population that speaks Spanish at home.

Which county has the highest percentage of 5-17 year-olds who speak Spanish at home and also speak English very well?

**A3**  Montgomery County, Maryland has the highest. I figured out how to join the counties but not the GEOID header. I arranged the results from the least to most, I kept getting an error when I wanted to do it from descending order. Luckily, the data isn't as big and I could scroll to the bottom to see.

```{r}
#Start with the code you wrote for Q1 and alter it to show the total population of 5-17 year-olds by county
mdcounties5_17<- get_acs(geography = "county", variables= "B16004_002", year= 2020, state= 'MD')  %>% 
  rename(total_pop = estimate) %>% 
  select(-variable) 


mdcounties_engverywell<-get_acs(geography = "county", variables= "B16004_005", year= 2020, state= 'MD') %>% 
rename(speak_well = estimate) %>% 
  select(-variable)



```

```{r}
together<- mdcounties5_17 %>% inner_join(mdcounties_engverywell, by=c("GEOID"="GEOID", "NAME"="NAME")) %>% 
arrange(desc(speak_well))

        

```  

**Q4** Given the counties with the largest percentage of 5-17 year-olds that speak English at home, which county stands out to you in the data as different from the others? How would you describe the top 5 counties? How would you describe Garrett County's numbers?

**A4** 
All of these counties have a large amount of adolescents that are in a suburban community.Being in Prince George's county, there's a lot of schools and a big focus on their education system and a curfew put in place for teenagers.The larger the population, the more adolescents that can speak English at home. Garrett County has a small number of adolescents and is not suburban, it looks very rural. 

**Q5** Pennsylvanians are currently requesting and submitting mail ballots for the November election. You'd like to know more about the people who live in the areas sending in mail ballots. Read in the `zip_totals.csv` file in the data folder containing the number of returned mail ballots per zip code. Then create a dataframe of Pennsylvania ZCTAs that include the following variables from the 2019 ACS: total population and median income. You will need to add `output = "wide"` to the get_acs function so that you get columns for those two variables' estimates and margins of error.

Join the ZCTA dataframe to the zip totals dataframe and calculate a per capita rate for mail ballots per 100 population in each ZCTA, and write code to answer the following questions?

* Which zip code *that has returned at least 100 mail ballots* has the highest per capita rate? 
* Where is that zip code and what does Census Reporter say about its demographics? 
* For zip codes with a per capita rate of at least 10 returned ballots per 100 people, how many have median incomes of at least $100,000? 
* How many have median incomes less than the state figure of $68,957?
* What's the most interesting one from this last result?

**A5**
Zipcode 17233 has the highest per capita rate. It is McConnelsburg County, Pennsylvania which has 1,150 people. Most of the peole living there are in their mid-40s and ninety-five percent white. So far, it's really rare to see someone who has a median income of $100,000. I see a trend of maybe the highest being $30,000 for the median income. I have a feeling this data is wrong but if it's true, the fact that there is little to none median incomes of at least 100,000 when at least 10 returned mail-in ballots. I think the McConnelsburg County is interesting if I calculated the per capita correctly, I highly doubt it, but I need to move on to Lab 5, 6, and now 8. I'll also review Lab 4, I want this lab to be the last Lab I lag behind. 

```{r}
zip_totals<-read.csv("data/pa_zip_totals.csv")
zip_totals <- zip_totals %>% mutate(across(zip,as.character))

#Then create a dataframe of Pennsylvania ZCTAs that include the following variables from the 2019 ACS: total population and median income.

acs1 <- load_variables(2019, "acs1") 
#Then create a dataframe of Pennsylvania ZCTAs that include the following variables from the 2019 ACS: total population and median income
mail_in_pop_medinc<- get_acs(output = "wide",geography = "zcta", variables= c(median_income="B07011_001", total_pop="B01003_001"), year= 2019, state= 'PA')

#Join the ZCTA dataframe to the zip totals dataframe 
joined_zip_pa<- mail_in_pop_medinc %>% inner_join(zip_totals, by=c("GEOID"="zip")) %>% 
 select(-NAME) %>% 
mutate(per_capita_rate=(total_popE/100)/count) %>% 
arrange(desc(per_capita_rate))
```
```{r}
  
```